## Getting and Cleaning Data Project
This is the submission Read Me for the Getting and Cleaning Data course project. This is the markdown version of the comments found in the run_analysis.R script.
  
The script accomplishes 5 things:
1) Merges the training and the test sets to create one data set.
2) Extracts only the measurements on the mean and standard deviation for each measurement. 
3) Uses descriptive activity names to name the activities in the data set
4) Appropriately labels the data set with descriptive variable names. 
5) From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.
  
Only one required library. The script will fetch and unzip the data to a 'data' subfolder in your working directory.
``` {r fetchandunzip}
library(RCurl)

if (!file.exists("./data")) {dir.create("./data")}
  fileURL <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
  download.file(fileURL, destfile="./data/UCI-HAR-dataset.zip", method = "curl")
  unzip("./data/UCI-HAR-dataset.zip", exdir = "./data")
```
  
Next we need to combine the training and test sets. They are unintuitively marked by default. Observations are Xs, labels are Ys, and subjects are factor coded only.
  
First, combine observations:
``` {r combineobs}
obs.train <- read.table("./data/UCI HAR Dataset/train/X_train.txt")
obs.test <- read.table("./data/UCI HAR Dataset/test/X_test.txt")
obs.data <- rbind(obs.train, obs.test)
```
Then, labels:
``` {r combinelabels}
label.train <- read.table("./data/UCI HAR Dataset/train/y_train.txt")
label.test <- read.table("./data/UCI HAR Dataset/test/y_test.txt")
label.data <- rbind(label.train, label.test)
```
Finally, subjects:
``` {r combinesubjects}
subj.train <- read.table("./data/UCI HAR Dataset/train/subject_train.txt")
subj.test <- read.table("./data/UCI HAR Dataset/test/subject_test.txt")
subj <- rbind(subj.train,subj.test)
```
Per the assignment instructions, we want only the standard deviation and mean for each measurement:
``` {r meanandsd}
desired.features <- read.table("./data/UCI HAR Dataset/features.txt")
mean.sd.features <- grep("-mean\\(\\)|-std\\(\\)", desired.features[,2])
x.mean.sd <- obs.data[, mean.sd.features]
```
We ought to rename the measurement names something useful and readable, extract the names from the features codebook:
``` {r nameextraction}
names(x.mean.sd) <- desired.features[mean.sd.features, 2]
names(x.mean.sd) <- tolower(names(x.mean.sd))
names(x.mean.sd) <- gsub("\\(|\\)", "", names(x.mean.sd))
```
Repeat a similar process for the activity names:
``` {r activityextraction}
activities <- read.table("./data/UCI HAR Dataset/activity_labels.txt")
activities[,2] <- tolower(as.character(activities[,2]))
activities[,2] <- gsub("_", "", activities[,2])
```
Now we rename the original labels per activity names and subjects.
``` {r renamelabels}
label.data[, 1] <- activities[label.data[,1], 2]
colnames(label.data) <- "activity"
colnames(subj) <- "subject"
```
Nearly done, reconstruct the merged set with the proper names.
``` {r mergeset}
data <- cbind(subj, x.mean.sd, label.data)
write.table(data, "./data/UCI HAR Dataset/merged_data.txt", row.names = FALSE)
str(data)
```
Finally, write out the smaller tidy set that contains the average measurement for each activity and each subject.
``` {r tidysetout}
average.set <- aggregate(x=data, by=list(activities=data$activity, subj=data$subject), FUN=mean)
average.set <- average.set[, !(colnames(average.set) %in% c("subj", "activity"))]
write.table(average.set, "./data/UCI HAR Dataset/average.txt", row.names = FALSE)
str(average.set)
```

